Hydras \& Co.

### 2022-03-16 doc/part-hydras.tex
corrections chapter 3

We  follow the same pattern as in Sect.~\ref{omega-case}.
First, we define an injection $\iota$ from type \texttt{t} into \texttt{Hydra}, by
 associating to  each ordinal $\omega\times i+ j = (i,j)$ the hydra with $i$ branches of length $2$ and
$j$ branches of length $1$.

@@

Like in Sect.~\ref{omega-case}, we prove the double inequality \texttt{m big\_h o<= m small\_h o< m big\_h}, which is impossible.

@@

The first steps of our proof prepare a well-founded induction on $\omega^2$.

@@

Then a case analysis on $i$ and $j$ allows us to
consider three cases :

\begin{itemize}
\item $i=j=0$: the inequality is trivial.
\item $i=1+l, j=0$ ($(i,j)$ is a limit ordinal): By the induction hypothesis \texttt{IHij},
  $(l,k)\leq m(\iota(l,k))$ for any $k$. But (by the rules of the hydra game), $\iota(i,0)$ is transformed into any $\iota(l,k)$ in one round. Thus $m(\iota(l,k)) < m(\iota(i,0))$ for any $k$.
  Therefore, $(l,k) <  m(\iota(i,0))$ for any $k$, thus
  $(i,0) \leq m(\iota(i,0))$.
 \item $j= l+1$  ($(i,j)$ is a successor).  a similar, but simpler case: we apply the induction hypothesis to the pair $(i,l)$.
 \end{itemize}

@@

\subsubsection{End of the proof}
From \texttt{m\_ge}, we get \texttt{m big\_h o<= m small\_h = m (iota (m big\_h)) }. 
Since $<$ is a strict order (irreflexive  and transitive), this inequality is incompatible with the strict inequality  \texttt{m small\_h o< m big\_h} (lemma \texttt{m\_lt}).

@@

\begin{remark}
In Chapter~\ref{ks-chapter}, we  prove a generalization of the impossibility lemmas of
Sect.~\ref{omega-case} and this section, with the same proof structure, but with much more 
complex technical details.
 \end{remark}

@@

\begin{remark}
Finite ordinals are also formalized in MathComp~\cite{MCB}.  See also Adam Chlipala's \emph{CPDT}~\cite{chlipalacpdt2011} for a thorough study of the use of dependent types.  
\end{remark}

@@

@@

Prove formally  that \texttt{ON\_mult (FinOrd $i$) Omega} is isomorphic to
\texttt{Omega}  whilst
\texttt{Omega}  is a sub-notation of \texttt{ON\_mult Omega (FinOrd $i$)},
for any strictly positive $i$.

>>>>

We  follow the same pattern as in Sect.~\ref{omega-case}.
First, we define an injection $\iota$ from type \texttt{ON\_Omega2.t} into \texttt{Hydra}, by
 associating to  each ordinal $\omega\times i+ j = (i,j)$ the hydra with $i$ branches of length $2$ and
$j$ branches of length $1$.

@@

Like in Sect.~\ref{omega-case}, we prove the inequality $m\;\texttt{big\_h} \,\texttt{o<=}\, m\;\texttt{small\_h} \,\texttt{o<=}\, m\;\texttt{big\_h}$, which is obviously false.

@@

The first steps of our proof start a well-founded induction over $\omega^2$.

@@

A case analysis on $i$ and $j$ allows us to
consider three cases :

\begin{itemize}
\item $i=j=0$: the inequality is trivial.
\item $i=1+l, j=0$ ($(i,j)$ is a limit ordinal): By the induction hypothesis \texttt{IHij},
  $(l,k)\texttt{o<=} m(\iota(l,k))$ for any $k$. But (by the rules of the hydra game), $\iota(i,0)$ is transformed into any $\iota(l,k)$ in one round. Thus $m(\iota(l,k)) < m(\iota(i,0))$ for any $k$.
  Therefore, $(l,k) <  m(\iota(i,0))$ for any $k$, thus
  $(i,0)\,\texttt{o<=}\,m(\iota(i,0))$.
 \item $j= l+1$  ($(i,j)$ is a successor). We apply the induction hypothesis to the pair $(i,l)$.
 \end{itemize}

@@

\subsubsection{End of the proof}
From \texttt{m\_ge}, we get \texttt{$m$\, big\_h o<= $m$\, small\_h = $m$\,(iota (m big\_h)) }. 
Since $<$ is a strict order (irreflexive  and transitive), this inequality is incompatible with the strict inequality  \texttt{m small\_h o< m big\_h} (lemma \texttt{m\_lt}).

@@

\begin{remark}
 We  prove in Chapter~\ref{ks-chapter} a generalization of the impossibility lemmas of
Sect.~\ref{omega-case} and this section, with the same proof structure, but with much more 
complex technical details.
 \end{remark}

@@

\section{See also \dots}

Finite ordinals are also formalized in \ssreflect/\mathcomp~\cite{MCB}. In Module~\href{../theories/html/gaia_hydras.ON_gfinite.html}{gaia\_hydras.ON\_gfinite}, we build an instance of class \texttt{ON}.

\inputsnippets{ON_gfinite/finordLtDef, ON_gfinite/finordON}

\inputsnippets{ON_gfinite/Examples, ON_gfinite/Examplesb}

  See also Adam Chlipala's \emph{CPDT}~\cite{chlipalacpdt2011} for a thorough study of the use of dependent types.

@@

\begin{project}
Same work, but replace Schütte's model with \gaia's.
\end{project}

@@

Prove formally  that (\texttt{ON\_mult (FinOrd $i$) Omega}) is isomorphic to
\texttt{Omega}  whilst
\texttt{Omega}  is a sub-notation of (\texttt{ON\_mult Omega (FinOrd $i$)}),
for any strictly positive $i$.
